Combinatorics of the geometry of Wilson loop diagrams I: equivalence classes via matroids and polytopes

### 2019-03-25 GWLDCI.tex
split into two papers

\title{Combinatorics of the geometry of Wilson loop diagrams}
\author{Susama Agarwala, Si\^an Fryer, and Karen Yeats}
%\date{}

@@

This paper studies the combinatorics of Wilson Loop Diagrams.
\section{Wilson Loop diagrams}\label{section background}

What are Wilson loop diagrams and their integrals.

@@

Generally speaking, the propagators are undirected, so $p = (i,j) = (j,i)$. In order to fix a convention, we write $p = (i,j)$ with  $i +1 < j$ relative to the first vertex. However, at times (section \ref{sec:propagator configs}), we consider directed propagators, where $p= (i,j)$ denotes a particular propagator flowing in one direction, while $p= (j,i)$ denotes the \emph{same} propagators flowing in the opposite direction. \sanote{Addendum 1}

@@

These $C(W)$ parametrize a subspace of $\Gr(k, n)$ as shown in \cite{wilsonloops}, call it $\Sigma(W)$. The Wilson loop diagrams also define a volume form on $\Sigma(W)$: \bas \Omega(W) = \frac{\prod_{r=1}^{|\cP|} \prod_{v \in V_{p_r}} \textrm{d}c_{p_r}}{R(W)} \;. \eas The denominator $R(W)$ is a polynomial defined by $2 \times 2$ and $1 \times 1$ minors of $C(W)$ as defined below.

\begin{dfn}\label{def R(W)}
For $W = (\cP, [n])$, $R(W) = \prod_{e=1}^n R_e$, with $R_e$ defined by the propagators ending on it. For any edge $e$ of $W$, order the propagators incident on $e$ as $\{p_1 \ldots p_r\}$, ordered such that $p_1$ is closest to the vertex $e$, $p_r$ closest to $e+1$, and $p_i$ is closer to $e$ than $p_{i+1}$. Then \bas R_e =  c_{p_1,e+1} \prod_{j= 1}^{r-1} \left((c_{p_j,e} c_{p_{j+1},e+1} - c_{p_{j+1},e} c_{p_{j},e+1} ) \right) c_{p_s,e}\;.\eas Note that in this notation, if $r = 1$, $R_e = c_{p,e} c_{p,e+1}$.
\end{dfn}

\section{Equivalence classes of Wilson loop diagrams}

In \cite{wilsonloop}, Agarwala and Amat show that Wilson loop diagrams can be interpreted as positroids, a certain well behaved class of realizable matroids (this correspondence is stated precisely in Theorem \ref{thm WLD defines matroid} below). This opens up the study of Wilson loop diagrams to techniques from geometry and combinatorics. In particular we examine the partial result shown in \cite{wilsonloop}, that shows that equivalent Wilson Loop Diagrams define the same matroid. In section 2.4, we make the relationship between Wilson Loop diagrams and matroids explicit. We count the size of each equivalence class, and enumerate the {\color{violet} the number of said classes}. We also show that two Wilson Loop diagrams map to the same matroid if and only if they are equivalent.

In subsection \ref{sec matroid background}, we discuss a few matroidal facts about Wilson loop diagrams. In section \ref{sec: polygon partitions}, we show that there is a one to one correspondence between exact subdiagrams of Wilson loop diagrams and triangulated pieces of the corresponding polygon partion. In section \ref{sec: exact diagram matroidal props}, we prove some matroidal properties of exact subdiagrams. We show that a subdiagram of $W$ defines a uniform matroid if and only if the subdiagram is exact (Theorem \ref{exactuniformthm}). Furthermore, this uniform matroid is the restriction of the dual matriod of $W$ restricted to the exact subdiagram (Remark \ref{remark exact dual restiction}). The main result of section \ref{sec: matroids and equivalence} shows that that two admissible Wilson loop diagrams define the same matroid if and only if they are equivalent (Theorem \ref{same matroid iff equiv}). We obtain a formula for the number of admissible Wilson loop diagrams in each equivalence class (Corollary \ref{number of equiv diagrams}) {\color{violet}as well as the number of equivalence classes}. In this way, we give a comple characterization of the relationship between Wilson Loop diagrams and their associated matroids.

@@

We first give a quick summary of the matroid terminology that we will need; it is not intended as a comprehensive introduction to matroids and the interested reader is referred to [ref][find a good matroid reference].

@@

\subsection{Polygon partitions of Wilson loop diagrams\label{sec: polygon partitions}}

The equivalence relation on Wilson loop diagrams is defined in terms of exact subdiagrams; thus in order to understand the equivalence, we need a way to extract and compare exact subdiagrams. We do this via the notion of a polygon partition of $W$.

\begin{dfn}
  Let $W = (\cP, [n])$ be an admissible Wilson loop diagram.  The \emph{polygon partition} associated to $W$, denoted $\tau(W)$, is defined as follows.
  \begin{itemize}
  \item The vertices of $\tau(W)$ correspond to the edges of $W$.
  \item Labeling the vertices of $\tau(W)$ with the edge number of $W$gives a cyclic order to the vertices. Connecting consecutive vertices gives a graph theoretic cycle called the polygon of $\tau(W)$.
  \item Each propagator of $W$ defines a chord edge of $\tau(W)$; specifically,  a propagator $(i,j) \in \cP$ defines a chord connecting the vertices $i$ and $j$ in $\tau(W)$.
  \end{itemize}
\end{dfn}

@@

\begin{comment}
\hlfix{For an admissible $W$ no propagators cross and so no chord edges of $\tau(W)$ cross.  Thus, in graph theoretic language, we can view $\tau(W)$ as a planar embedding of a
graph with no cut vertices, where all vertices lie on the infinite face, along with a distinguished start vertex and direction around the infinite face.  From this viewpoint the polygon of $tau(W)$ is the boundary of the infinite face.}{I've smushed this together a bit in the prev. lem. Tell me if I've missed anything. Its a lemma for emphasis, not difficulty}
\end{comment}

\begin{eg}\label{WLDtopolygonpartition}
In this example we return to two of the Wilson loop diagrams in Example \ref{eg:equivdiags}. We can pair diagrams with their polygon partitions as follows:
\bas W_1 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
	\drawprop{1}{0}{4}{0}
	\drawprop{2}{0}{4}{-1}
    \drawprop{5}{0}{8}{0}
    \drawprop{5}{1}{7}{0}
		\end{scope}
	\end{tikzpicture} \quad; \quad
\tau(W_1) = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawpolypart{8}{1.5}
    \drawnumbersshift
    \drawchord{1}{4}
    \drawchord{2}{4}
    \drawchord{5}{8}
    \drawchord{5}{7}
	\end{scope}
	\end{tikzpicture}
\eas and
\bas W_3 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
    \drawprop{1}{0}{4}{0}
	\drawprop{1}{1}{3}{0}
    \drawprop{5}{0}{8}{0}
    \drawprop{6}{0}{8}{-1}	
		\end{scope}
	\end{tikzpicture}\quad; \quad
\tau(W_3) = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawpolypart{8}{1.5}
    \drawnumbersshift
    \drawchord{1}{4}
    \drawchord{1}{3}
    \drawchord{5}{8}
    \drawchord{6}{8}
	\end{scope}
	\end{tikzpicture} .
\eas

@@

\begin{dfn}
  Let $W$ be an admissible Wilson loop diagram and $\tau(W)$ its polygon partition. A \emph{triangulated piece} of $\tau(W)$ is a 2-connected subgraph of $\tau(W)$ which is a triangulation. We will take the convention that a subgraph consisting of a single chord edge is called a \emph{trivial} triangulated piece.
A {\em maximal} triangulated piece is one which is not contained in any strictly larger triangulated piece.
\end{dfn}

\begin{dfn}
 A {\em decomposition} of a polygon partition $\tau(W)$ is a set of 2-connected induced subgraphs of $\tau(W)$ which partition the edges of $\tau(W)$.  
\end{dfn}

\begin{eg} \label{eg: unique decomposition} For the Wilson loop diagrams and polygon partitions in Example \ref{WLDtopolygonpartition}, the vertex sets $\{1, 2, 3,4\}$ and $\{5, 6, 7, 8\}$ give maximal triangulated pieces for both $\tau(W_1)$ and $\tau(W_3)$. The vertex set $\{4,5, 8, 1\}$ is not a triangulation in either polygon partition. 
\end{eg}

\begin{lem} \label{decompositionlem}
  For $W$ an admissible Wilson loop diagram, the polygon partition $\tau(W)$ has a unique decomposition into maximal triangulated pieces, and edges in the polygon of $\tau(W)$.
\end{lem}

@@

\begin{lem}\label{lem triang to exact}
  Let $W$ be an admissible Wilson loop diagram and $\tau(W)$ its polygon partition.  The triangulated pieces of $\tau(W)$ correspond to the exact subdiagrams of $W$ using the correspondence described above.
\end{lem}

@@

\subsection{Matroids and equivalent diagrams \label{sec: matroids and equivalence}}
\note{Si\^an postponed reading this section until the earlier ones are fixed.}
Now we are ready to prove the main results of this section, namely that two Wilson loop diagrams define the same matroid if and only if they are equivalent (Theorem \ref{same matroid iff equiv}). We also {\color{violet}count the number of equivalence classes amongst Wilon Loop diagrams with $n$ vertices and $k$ propagators and} give a formula for the number of Wilson loop diagrams in a given equivalence class (Corollary \ref{number of equiv diagrams}), completing the characterization of the correspondence between Wilson Loop diagrams and matroids starting in \cite{wilsonloop}.

@@

Since there is a unique way to decompose $W$ into maximal exact subdiagrams, it is logical to ask how many diagrams there are in an equivalence class. It is a classical fact the the number of triangulations of an $n$-gon is the $n-2$ Catalan number, namely $\frac{1}{n-1}\binom{2(n-2)}{n-2}$.  Thus we can count the number of equivalent diagrams. [CITATION?]

@@

In this section we have characterized the correspondence between Wilson loop diagrams and positroids, showing that Wilson Loop diagrams define the same matroid if and only if they are equivalent (Theorem \ref{same matroid iff equiv}). Corollary \ref{number of equiv diagrams} enumerates the fibres of the map from Wilson Loop diagrams to positroids. {\color{violet} By counting the number of equivalence classes (Theorem \ref{}), we enumerate the number of positroids defined by Wilson Loop diagrams. In section \ref{}, Theorem \ref{}, we show that each Wilson Loop diagram with $n$ vertices and $k$ propagators define a positroid cell of dimension $3k$. Comparing the result of Theorem \ref{} to the number of positroid cells of a particular dimension in $\Grtnn(n, k)$, [CITE AND FIND FORMULA HERE], we see that \emph{Wilson loop diagrams do not map onto all positroids of the correct dimension.}}

\begin{comment}
\subsection{Counting exact subdiagrams}

\begin{dfn}
  Say two admissible Wilson loop diagrams $W_1$ and $W_2$ are \emph{triangulation-equivalent} if there is a bijection $\alpha$ between the set of triangulated pieces in the decomposition of $\tau(W_1)$ and the set of triangulated pieces in the decomposition of $\tau(W_2)$, where $t$ and $\alpha(t)$ have the same vertex set for all triangulated pieces $t$ of $\tau(W_1)$.
\end{dfn}

***draw an example***

\begin{thm}
  Two admissible Wilson loop diagrams have the same matroid iff they are triangulation equivalent
\end{thm}

Again do I want matroid or positroid in this theorem?

\begin{proof}
  By Lemma~\ref{lem triang to exact} two admissible Wilson loop diagrams are triangulation equivalent iff their decompositions into maximal connected exact subdiagrams have matching supports, that is if they are equivalent.
  Theorem~\ref{thm orig equiv} then gives the result.
\end{proof}

\section{Geometry of Wilson Loop diagram}\label{sec GN algorithm}

Since Wilson loop diagrams correspond to positroids, it is natural to study the subspace of $\Gr(|\cP|, n)$ they define.

[words about what's in this section]

Throughout this section, $W = (\cP,[n])$ is an admissible Wilson loop diagram with $k$ propagators.

\subsection{Background}\label{sec:positroid background}

[positroids]

Let $\binom{[n]}{k}$ be the set of all $k$-subsets of the cyclically ordered set $[n]$.  For each $j \in [n]$, we can define a total order $\leq_j$ on the interval $[n]$ by
\[ j <_j j+1 <_j \dots <_j n <_j 1 \dots <_j j-1\;.\]
This in turn induces a total order on $\binom{[n]}{k}$, namely the lexicographic order with respect to $<_j$.  It also induces a separate partial order $\gale{j}$ on $\binom{[n]}{k}$ (the {\bf Gale order} \cite{Gale}), which is defined as follows: for 
\[A = \{a_1 <_j a_2 <_j \dots <_j a_k\} \text{ and } B = \{b_1 <_j b_2 <_j \dots <_j b_k\} \in \binom{[n]}{k},\] we define
\[A \gale{j} B \text{ if and only if } a_r \leq_j b_r \text{ for all }1 \leq r \leq k.\]
For example, in $\binom{[6]}{3}$ we have $\{2,5,6\}\gale{2} \{2,6,1\}$ but $\{2,5,6\}\not\gale{2}\{3,4,6\}$.

\begin{dfn}\label{def:grassmann necklace}
A {\bf Grassmann necklace} of type $(k,n)$ is a sequence $(I_1, \dots, I_n)$ of $n$ sets $I_i \in \binom{[n]}{k}$ such that for each $i \in [n]$:
\begin{itemize}
\item if $i \in I_i$, then $I_{i+1} = \big(I_i \backslash \{i\}\big) \cup \{j\}$ for some $j \in[n]$.
\item if $i \not\in I_i$, then $I_{i+1} = I_i$.
\end{itemize}
By convention, we set $I_{n+1} = I_1$.
\end{dfn}

By \cite[Theorem 17.1]{Postnikov}, the Grassmann necklaces of type $(k,n)$ are in 1-1 correspondence with the positroid cells in $\Gr(k,n)$.  This correspondence is given explicitly in \cite[Theorem 8]{Oh}: if $(I_1, \dots, I_n)$ is the Grassmann necklace associated to a positroid $M = ([n],\cB)$, then the bases of $M$ are exactly
\[\cB = \left\{J \in \binom{[n]}{k}\ :\ I_i \gale{i} J \ \forall i \in [n]\right\}.\]

\begin{dfn}\label{def:le diagram}
A {\bf Le diagram} is a Young diagram in which every square contains either a $+$ or a $0$, subject to the rule that if a square contains a $0$ then either all squares to its left (in the same row) must also contain a $0$, or all squares above it (in the same column) must also contain a $0$, or both.
\end{dfn}

By [ref], the set of all Le diagrams that fit within a $k\times(n-k)$ rectangle is in 1-1 correspondence with the positroid cells of $\Gr(k,n)$. The dimension of a positroid cell is equal to the number of $+$ squares in its Le diagram [ref].

The rows and columns of a Le diagram are labelled as follows: given a Le diagram fitting inside a $k\times (n-k)$ box, arrange the numbers $1,2, \dots, n$ along its southeast border, starting from the top-right corner. See Figure \ref{fig:row column numbering} for examples.

\begin{figure}[h!]
\[\begin{tikzpicture}[baseline=(current bounding box.east),scale=0.75]
\draw (1,0) grid (2,3);
\draw (2,1) grid (4,3);
\draw (4,2) grid (5,3);
\draw[dotted] (2,0) -- (5,0) -- (5,2);

\node at (5.15,2.5) {\footnotesize$1$};
\node at (4.5,1.85) {\footnotesize$2$};
\node at (4.15,1.5) {\footnotesize$3$};
\node at (3.5,0.85) {\footnotesize$4$};
\node at (2.5,0.85) {\footnotesize$5$};
\node at (2.15,0.5) {\footnotesize$6$};
\node at (1.5,-0.15) {\footnotesize$7$};
%\node at (0.5,-0.15) {\footnotesize$8$};
\end{tikzpicture}
\qquad \qquad
\begin{tikzpicture}[baseline=(current bounding box.east),scale=0.75]
\draw (1,0) grid (2,3);
\draw (2,1) grid (4,3);
\draw (4,2) grid (5,3);
\draw[dotted] (2,0) -- (6,0) -- (6,3) -- (5,3);

\node at (5.5,2.85) {\footnotesize$1$};
\node at (5.15,2.5) {\footnotesize$2$};
\node at (4.5,1.85) {\footnotesize$3$};
\node at (4.15,1.5) {\footnotesize$4$};
\node at (3.5,0.85) {\footnotesize$5$};
\node at (2.5,0.85) {\footnotesize$6$};
\node at (2.15,0.5) {\footnotesize$7$};
\node at (1.5,-0.15) {\footnotesize$8$};
%\node at (0.5,-0.15) {\footnotesize$9$};
\end{tikzpicture}
\]
\caption{Row and column numbering for a Young diagram with $k = 3$, $n = 7$ (left) and $k = 3$, $n = 8$ (right). The top left box in each diagram has coordinates $(1,7)$ (left diagram), $(2,8)$ (right diagram).}
\label{fig:row column numbering}
\end{figure}

An algorithm for constructing the Le diagram associated to a Grassmann necklace was given by Agarwala and Fryer in [ref]. Since we will make use of this algorithm in Section [ref] below, we summarise the process here.
\begin{algorithm}\label{alg:GN to Le} \ [ref]
Let $(I_1,\dots,I_n)$ be a Grassmann necklace of type $(k,n)$. Within a $k \times(n-k)$ square, draw the Young diagram whose rows are labelled by $I_1$ (as per the convention above).

>>>>

\title{Combinatorics of the geometry of Wilson loop diagrams I: equivalence classes via matroids and polytopes}
\author{Susama Agarwala, Si\^an Fryer, and Karen Yeats}
%\date{}

@@

\section{Introduction}

***Some general intro to WLDs perhaps similar to the end of p2 and top of p3 of \cite{wilsonloop}.  For this paper we don't need the integral, so for the moment they are not defined below, but for motivation may want to be glossed a bit here ***

Furthermore, in \cite{wilsonloop}, Agarwala and Amat show that Wilson loop diagrams can be interpreted as positroids, a certain well behaved class of realizable matroids.  This opens up the study of Wilson loop diagrams to techniques from geometry and combinatorics.  We pursue this direction here and in the companion paper ***cite the other paper***.

In the present paper we focus on the combinatorics and geometry of an equivalence relation on Wilson loop diagrams.  The main results are that Wilson loop diagrams are equivalent if and only if they define the same matroid, that the equivalence can be rephrased in terms of retriangulations in polygon dissections, and from there that the sizes of the equivalence classes can be enumerated, and the set of equivalence classes is related to directions in associahedra.

\subsection{Roadmap}

We proceed in Section~\ref{section overall background} with the required definitions for Wilson loop diagrams (subsection~\ref{section WLD background}) and matroids (subsection~\ref{sec matroid background}.  Wilson loop diagrams will be said to be admissible when they correspond to the physical situation of ***something***, or mathematically according to Definition~\ref{admisdfn}.  With this, the key definition is Definition~\ref{equivdfn} giving the equivalence relation on admissible Wilson loop diagrams.  The connection to matroids is stated precisely in Theorem \ref{thm WLD defines matroid}.

The main result of Section~\ref{sec equiv} is Theorem~\ref{same matroid iff equiv}.  To arrive at those results, we first show how certain subdiagrams, called exact subdiagrams, relate to triangulations in polygons in subsection~\ref{sec: polygon partitions}, and then consider the matroidal properties of the exact subdiagrams in subsection~\ref{sec: exact diagram matroidal props}.  Specifically, a subdiagram of a Wilson loop diagrams $W$ defines a uniform matroid if and only if the subdiagram is exact (Theorem \ref{exactuniformthm}), and this uniform matroid is the restriction of the dual matriod of $W$ restricted to the exact subdiagram (Remark \ref{remark exact dual restiction}).  The main result of the section, Theorem~\ref{same matroid iff equiv}, then appears in subsection~\ref{sec: matroids and equivalence}.  A consequence of this is the formula for the size of an equivalence class given in Corollary~\ref{number of equiv diagrams}.

The next natural question, then is the number of inequivalent diagrams under this equivalence relation.  The main result of Section~\ref{sec associahedron} is to show that the set of admissible Wilson loop diagrams up to equivalence is in bijection with the set of faces in the associahedron up to paralellness.  Towards this result we go over some relevant background on polytopes in subsection~\ref{sec polytope background}, and then give the main result in subsection~\ref{sec associahedron results} in the form of two propositions, Propositions~\ref{prop easy way} and~\ref{prop hard way}.

\section{Background}\label{section overall background}

\subsection{Wilson Loop diagrams}\label{section WLD background}

%What are Wilson loop diagrams and their integrals.

@@

Propagators are undirected, so $p = (i,j) = (j,i)$. In order to fix a convention, we write $p = (i,j)$ with  $i +1 < j$ relative to the first vertex.

@@

These $C(W)$ parametrize a subspace of $\Gr(k, n)$ as shown in \cite{wilsonloop}, call it $\Sigma(W)$.

%The Wilson loop diagrams also define a volume form on $\Sigma(W)$: \bas \Omega(W) = \frac{\prod_{r=1}^{|\cP|} \prod_{v \in V_{p_r}} \textrm{d}c_{p_r}}{R(W)} \;. \eas The denominator $R(W)$ is a polynomial defined by $2 \times 2$ and $1 \times 1$ minors of $C(W)$ as defined below.
%
%\begin{dfn}\label{def R(W)}
%For $W = (\cP, [n])$, $R(W) = \prod_{e=1}^n R_e$, with $R_e$ defined by the propagators ending on it. For any edge $e$ of $W$, order the propagators incident on $e$ as $\{p_1 \ldots p_r\}$, ordered such that $p_1$ is closest to the vertex $e$, $p_r$ closest to $e+1$, and $p_i$ is closer to $e$ than $p_{i+1}$. Then \bas R_e =  c_{p_1,e+1} \prod_{j= 1}^{r-1} \left((c_{p_j,e} c_{p_{j+1},e+1} - c_{p_{j+1},e} c_{p_{j},e+1} ) \right) c_{p_s,e}\;.\eas Note that in this notation, if $r = 1$, $R_e = c_{p,e} c_{p,e+1}$.
%\end{dfn}

@@

We first give a quick summary of the matroid terminology that we will need; it is not intended as a comprehensive introduction to matroids and the interested reader is referred to [ref][find a good matroid reference].\note{Oxley is always my go to reference for matroids, but perhaps it has the wrong focus for you?}

@@

\section{Matroids, triangulations, and equivalence classes of Wilson loop diagrams}\label{sec equiv}

\subsection{Polygon dissections of Wilson loop diagrams\label{sec: polygon partitions}}

The equivalence relation on Wilson loop diagrams is defined in terms of exact subdiagrams; thus in order to understand the equivalence, we need a way to extract and compare exact subdiagrams. We do this via the notion of a polygon dissection of $W$.

\begin{dfn}
  Let $W = (\cP, [n])$ be an admissible Wilson loop diagram.  The \emph{polygon dissection} associated to $W$, denoted $\tau(W)$, is defined as follows.
  \begin{itemize}
  \item The vertices of $\tau(W)$ correspond to the edges of $W$.
  \item Labeling the vertices of $\tau(W)$ with the edge number of $W$gives a cyclic order to the vertices. Connecting consecutive vertices gives a graph theoretic cycle called the polygon of $\tau(W)$.
  \item Each propagator of $W$ defines a chord edge of $\tau(W)$; specifically,  a propagator $(i,j) \in \cP$ defines a chord connecting the vertices $i$ and $j$ in $\tau(W)$.
  \end{itemize}
\end{dfn}

\emph{Polygon dissection} is a common term in combinatorics for a decomposition of a convect polygon into smaller polygons by means of noncrossing diagonals, see \cite{PSpolygon}.  Each $\tau(W)$ is such an object.

@@

\begin{eg}\label{WLDtopolygonpartition}
In this example we return to two of the Wilson loop diagrams in Example \ref{eg:equivdiags}. We can pair diagrams with their polygon dissections as follows:
\bas W_1 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
	\drawprop{1}{0}{4}{0}
	\drawprop{2}{0}{4}{-1}
    \drawprop{5}{0}{8}{0}
    \drawprop{5}{1}{7}{0}
		\end{scope}
	\end{tikzpicture} \quad; \quad
\tau(W_1) = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawpolypart{8}{1.5}
    \drawnumbersshift
    \drawchord{1}{4}
    \drawchord{2}{4}
    \drawchord{5}{8}
    \drawchord{5}{7}
	\end{scope}
	\end{tikzpicture}
\eas and
\bas W_3 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
    \drawprop{1}{0}{4}{0}
	\drawprop{1}{1}{3}{0}
    \drawprop{5}{0}{8}{0}
    \drawprop{6}{0}{8}{-1}	
		\end{scope}
	\end{tikzpicture}\quad; \quad
\tau(W_3) = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawpolypart{8}{1.5}
    \drawnumbersshift
    \drawchord{1}{4}
    \drawchord{1}{3}
    \drawchord{5}{8}
    \drawchord{6}{8}
	\end{scope}
	\end{tikzpicture} .
\eas

@@

\begin{dfn}
  Let $W$ be an admissible Wilson loop diagram and $\tau(W)$ its polygon dissection. A \emph{triangulated piece} of $\tau(W)$ is a 2-connected subgraph of $\tau(W)$ which is a triangulation. We will take the convention that a subgraph consisting of a single chord edge is called a \emph{trivial} triangulated piece.
A {\em maximal} triangulated piece is one which is not contained in any strictly larger triangulated piece.
\end{dfn}

\begin{dfn}
 A {\em decomposition} of a polygon dissection $\tau(W)$ is a set of 2-connected induced subgraphs of $\tau(W)$ which partition the edges of $\tau(W)$.  
\end{dfn}

\begin{eg} \label{eg: unique decomposition} For the Wilson loop diagrams and polygon dissections in Example \ref{WLDtopolygonpartition}, the vertex sets $\{1, 2, 3,4\}$ and $\{5, 6, 7, 8\}$ give maximal triangulated pieces for both $\tau(W_1)$ and $\tau(W_3)$. The vertex set $\{4,5, 8, 1\}$ is not a triangulation in either polygon dissection. 
\end{eg}

\begin{lem} \label{decompositionlem}
  For $W$ an admissible Wilson loop diagram, the polygon dissection $\tau(W)$ has a unique decomposition into maximal triangulated pieces, and edges in the polygon of $\tau(W)$.
\end{lem}

@@

\begin{lem}\label{lem triang to exact}
  Let $W$ be an admissible Wilson loop diagram and $\tau(W)$ its polygon dissection.  The triangulated pieces of $\tau(W)$ correspond to the exact subdiagrams of $W$ using the correspondence described above.
\end{lem}

@@

\subsection{Matroids and equivalent diagrams \label{sec: matroids and equivalence}}
\note{Si\^an postponed reading this section until the earlier ones are fixed.}
Now we are ready to prove the main results of this section, namely that two Wilson loop diagrams define the same matroid if and only if they are equivalent (Theorem \ref{same matroid iff equiv}).
We also
give a formula for the number of Wilson loop diagrams in a given equivalence class (Corollary \ref{number of equiv diagrams}), completing the characterization of the correspondence between Wilson Loop diagrams and matroids starting in \cite{wilsonloop}.

@@

Since there is a unique way to decompose $W$ into maximal exact subdiagrams, it is logical to ask how many diagrams there are in an equivalence class. It is a classical fact (see for instance Chapter 6 of \cite{Stanley}) that the number of triangulations of a convex $n$-gon is the $n-2$ Catalan number, namely $\frac{1}{n-1}\binom{2(n-2)}{n-2}$.  Thus we can count the number of equivalent diagrams.

@@

In this section we have characterized the correspondence between Wilson loop diagrams and positroids, showing that Wilson Loop diagrams define the same matroid if and only if they are equivalent (Theorem \ref{same matroid iff equiv}). Corollary \ref{number of equiv diagrams} enumerates the fibres of the map from Wilson Loop diagrams to positroids.  The obvious next question would be to enumerate the number of equivlence classes.  We do not obtain a formula for this, but do obtain an alternate characterization of the equvalence classes which is the subject of the next section.

\section{Associahedra and inequivalent Wilson loop diagrams}\label{sec associahedron}

The connection between admissible Wilson loop diagrams and polytope dissections suggests a connection to associahedra. First we need some background.

\subsection{Polytope background}\label{sec polytope background}

A \emph{polytope} can equivalently be defined as the convex hull of a finite set of points in some $\mathbb{R}^d$ or as the intersection of finitely many closed half spaces in $\mathbb{R}^d$ with the additional property that this intersection is bounded in the sense of not containing any ray (see chapter 0 of \cite{Ziegler}).

A \emph{face} of a polytope is the intersection of the polytope with any hyperplane for which the entire polytope is in one of the closed half-spaces defined by the hyperplane.  The polytope itself is also considered a face (in the case that that polytope is of full dimension, this definition can be unified by considering linear equalities (including the trivial one) in place of hyperplanes and the corresponding linear inequalities in place of the half-spaces defined by the hyperplane, see chapter 2 of \cite{Ziegler}).  Two faces are \emph{parallel} if one can be mapped to the other by a translation in $\mathbb{R}^d$.  In particular parallel faces must be of the same dimension and any pair of vertices are parallel.

There are many important polytopes in combinatorics, including, of primary importance to us, the associahedron.

Two polytopes are \emph{combinatorially equivalent} if there is a bijection between their faces which preserves inclusion of faces.  Often in combinatorics one is primarily interested in the combinatorial equivalence classes, and common names, including the associahedron, refer to these equivalence classes.  A particular polytope in the class is then known as a \emph{realization}.

Associahedra can also be viewed in terms of polygon dissections.  Let's fix $n$ and consider dissections of a convex $n$-gon.  The $n-1$ asociahedron can be given in terms of the dissections of this $n$-gon; the vertices of the associahedron are the triangulations of the $n$-gon while the dimension $k$ faces are the dissections which are $k$ diagonals away from being a triangulation.  One face is included in another if the first dissection is a refinement of the second.

It turns out that this defines a class of combinatorially equivalent polytopes, which can be proved by giving a realization.  There are many interesting realizations of the associahedron \cite{}; the following will be the one which we will use.

\begin{dfn}
  Let $x_1, \ldots, x_n$ define a convex $n$-gon.  Let $T$ be the set of triangulations of this $n$-gon.  For each $t\in T$ define $s_t$ to be the point in $\mathbb{R}^n$ whose $i$th coordinate is the sum of the areas of all triangles of $t$ which are incident to $x_i$.  Let $A_n$ be the convex hull of the $s_t$.
\end{dfn}
This is the construction of the secondary polytope (so $A_n$ is the secondary polytope of $x_1, \ldots, x_n$), see \cite{Ziegler} Definition 9.9.  In general the $x_1, \ldots, x_n$ may be any points in any $\mathbb{R}^d$.  Triangulations in the more general context are a decomposition into simplices and the areas of the triangles is replaced by the volume of the simplices.

$A_n$ defined in this way is a realization of the $n-1$ associahedron (\cite{Ziegler}, Example 9.11).  Note that we get a realization for any convex $n$-gon, but these realizations are all equally good for our purposes, and so we fix $x_1, \ldots, x_n$ from now on.

\subsection{Non-parallel faces in the associahedron and inequivalent Wilson loop diagrams}\label{sec associahedron results}

With the choice of $n$-gon and hence of $A_n$ fixed, view the $\tau$ map as taking values in the set of dissections of this particular convex $n$-gon, and so each weakly admissible Wilson loop diagram is associated via $\tau$ with a face (in some dimension) of $A_n$.

Note that Wilson loop diagrams which are weakly admissible but not admissible have $\tau$ graphs which are triangulations, hence dimension 0 faces (vertices) in $A_n$.  All of these are equivalent in the sense that they are only retriangulations away from each other.  It will be convenient in what follows, to include these diagrams.  The next two propositions would be identical in the case of admissible instead of weakly admissible, while the corollary could be changed to involve admissible diagrams only by subtracting one from the number of non parallel faces in order to account for the missing dimension 0 faces.

\begin{prop}\label{prop easy way}
  If two weakly admissible Wilson loop diagrams are equivalent then their faces in $A_n$ are parallel.
\end{prop}

\begin{proof}
  Let $W$ and $X$ be equivalent.  From results in the paper we're working on we know that $\tau(W)$ and $\tau(X)$ differ from each other only by retriangulations of their triangulated pieces.

  All edges of $A_n$ are polygon dissections which are only one diagonal away from being triangulations.  Such dissections consist of triangles and exactly one quadrilateral.  Consider the edges of $A_n$ bounding the faces corresponding to $\tau(W)$ and $\tau(X)$.  These edges have their one quadrilateral within the nontriangulated parts of $\tau(W)$ or $\tau(X)$, but these nontriangulated parts are the same in $\tau(W)$ and $\tau(X)$, so we get a bijection between the edges bounding the face of $\tau(W)$ and the edges bounding the face of $\tau(X)$ which matches edges with the same quadrilateral.

  Consider now an edge whose quadrilateral has corners $x_i$, $x_j$, $x_k$, and $x_l$ in that order.  Let $t_1$ and $t_2$ be the triangulations corresponding to the two ends of the edge where $t_1$ has a diagonal from $x_j$ to $x_l$ and $t_2$ has a diagonal from $x_i$ to $x_k$.  From the definition of $A_n$ we see that $s_{t_1}$ and $s_{t_2}$ only differ at coordinates $i$, $j$, $k$, and $l$, and all of the triangles incident to any of these vertices other than the one coming from triangulating $x_i, x_j, x_k, x_l$ are the same in both $t_1$ and $t_2$.  Let $a$ be the area of the quadrilateral and let $a(i,j,k)$ be the area of the triangle with corners $x_i$, $x_j$ and $x_k$, and similarly for other indices then $s_{t_1}$ on coordinates $i$, $j$, $k$, $l$ is $(a(i,j,l) + b_i, a+b_j, a(k,l,j)+b_k, a+b_l)$ and $s_{t_2}$ on these coordinates is $(a + b_i, a(j,k,i)+b_j, a+b_k, a(l,i,k)+b_l$, where the $b_i$ are the sums of the areas of the relevant triangles which are common to both triangulations.  Thus a vector parallel to this edge is the vector which is $(a(i,j,l) - a, a-a(j,k,i), a(k,l,j)-a, a-a(l,i,k)) = (-a(j,k,l), a(i,k,l), -a(i,j,l), a(i,j,k))$ on $i$, $j$, $k$, $l$ and $0$ in all other coordinates.

  Returning to $\tau(W)$ and $\tau(X)$ since the direction vector of an edge in $A_n$ depends only on its quadrilateral and the bijection between the edges bounding $\tau(W)$ and the edges bounding $\tau(X)$ preserves the quadrilateral, we have the same direction vectors bounding $\tau(W)$ as $\tau(X)$.  Furthermore, the bijection also maintains the relative configuration of the edges and so $\tau(W)$ and $\tau(X)$ are parallel.
\end{proof}

The following lemma is useful for the second proposition.
